Viscous Transport in Eroding Porous Media

### 2019-06-07 docs/transport/erosion_transport.tex
worked in section 4.3

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fast Summation Methods}
\label{sec:fmm}
The Barycentric quadrature rule achieves acceptable accuracy for
computing the velocity, shear stress, and vorticity.  However, applied
directly, it requires $\mathcal{O}(N^2)$ operations.  Since the shear
stress must only be computed once per time step, and the vorticity is
only used as a post-processing step, we accept the $\mathcal{O}(N^2)$
algorithm.  However, the velocity must be computed at every GMRES
iteration, which would make the overall method impractical without a
fast algorithm.

@@

We also do not use any fast summation methods to apply the Barycentric
quadrature rule.  Since this quadrature rule is applied at every GMRES
iteration, it is imperative that we only apply the Barycentric
quadrature at target points where they are required.  As a rule of
thumb, the trapezoid rule due to $\gamma_k$ is accurate if
\begin{align}
  d(\xx,\gamma_k) = \inf_{\yy \in \gamma_k} \|\xx - \yy\| > 
    5 \frac{L_k}{N_\iin}.
\end{align}
To avoid having to check if a target point $\xx$ satisifies

The same strategy of using the trapezoid rule and then correcting with
the Barycentric quadrature is used to compute the shear stress on
$\bd\Omega$ and vorticity in $\Omega$.  The trapezoid rule is applied
with a direct summation, but this cost is only incurred once per time
step, and the majority of the computational effort is still required to
compute the velocity using the fast summation method.\todo[inline]{BQ:
Discuss how we don't do Barycentric everywhere since it involves several
$N$ body calculations, while the trapezoid rule only requires a single
one}

To determine if a point $\xx \in \gamma_i$ needs to be corrected with
the Barycentric quadrature rule, we first estimate the proximity of
$\xx$ to $\bd\Omega$.  The criteria depends on the length of the outer
wall, $L_w$, the length of the eroding bodies $L_i$, $i=1,\ldots,M$, and
the number of discretization points $N_\out$ and $N_\iin$.

{\color{red}
By using Barycentric quadrature rule to approach integrations, we get a reasonable(?) double-layer potential and quantities of interest at target points but it requires $\mathcal{O}(N^2)$ operations per target point which is computational expensive if we impose it on all the target points. In order to accelerate the process and keep the accuracy, it is a compromise that we use Barycentric quadrature rule to approach the nearly singular integratons but apply trapezoid rule when the target point is far from source points.

To numerically compute the completed double-layer potential, we discretize each of the $M$ bodies with  $N_\iin$ points, and the outer wall with $N_\out$ points. The following is the four-step fast summation methods.
}
\begin{itemize}
  \item Apply trapezoid rule as a first pass

{\color{red} We use the Fast Multipole Method (FMM)~\cite{gre-rok1987, gre-gre-may1992} with trapezoid rule to calculate the double-layer potential.
}
  \item Efficiently find points that are too close for trapezoid

{\color{red}
 Let $L_w$ be the length of the outer wall and $L_{b}$ be the length of the boundary of body where $b=1, \cdots, M$. Here are the criteria to see if two points are too close for trapezoid. For criterion of outer wall to body,  let $d$ be the minimum distance between outer wall and a point $x$ on the body $b$, we say $x$ is too close to outer wall if
\begin{equation}\label{eqn:Cw2b}
d< \alpha_1\frac{L_w}{N_\out}
\end{equation}
where $\alpha_1$ is a constant. 
For criterion between two bodies, let $d$ be the distance between the centroid $a$ of the body $b_1$ and a point $x$ on the body $b_2$, we say they are too close if
\begin{equation}\label{eqn:Cb2b}
d<\alpha_2\left(\frac{L_{b_1}}{N_\iin}+\frac{L_{b_2}}{N_\iin}\right)
\end{equation}
where $\alpha_2$ is a constant.

}
\end{itemize}
{\color{red}  (Discussion of the choices of $\alpha_1$ and $\alpha_2$ with respect to accuracy and  efficiency)

To find the suitable values of $\alpha_1$ and $\alpha_2$ in~\eqref{eqn:Cw2b} and \eqref{eqn:Cb2b}, we set $\alpha_1=\alpha_2=1, 2, 3$, and $4$ to test the criteria on a 20-body problem with $N_\out = 1024$ and $N_\iin = 128, 256$, and $512$. To check the accuracy of the simulations, we use the solution calculated by \todo[inline]{what method} to \todo[inline]{...}
To see how efficient our code is, we record the cpu time for each step.

}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Efficiently Applying Quadrature}
\label{sec:fmm}
The Barycentric quadrature rule achieves acceptable accuracy for
computing the velocity, shear stress, and vorticity.  However, applied
directly, it requires $\mathcal{O}(N^2)$ operations.  Since the shear
stress must only be computed once per time step, and the vorticity is
only used as a post-processing step, we accept the $\mathcal{O}(N^2)$
algorithm.  However, the velocity must be computed at every GMRES
iteration, which would make the overall method impractical without a
fast algorithm.

@@

We also do not use any fast summation methods to apply the Barycentric
quadrature rule.  Since this quadrature rule is applied at every GMRES
iteration, it is imperative that we only apply the Barycentric
quadrature at target points where they are required.  As a rule of
thumb, the trapezoid rule due to $\gamma_k$ is accurate if
\begin{align}
  \label{eqn:TrapCutoff}
  d(\xx,\gamma_k) = \inf_{\yy \in \gamma_k} \|\xx - \yy\| > 
    5 \frac{L_k}{N_\iin}.
\end{align}
To avoid checking if a target point $\xx$
satisfies~\eqref{eqn:TrapCutoff}, we make use of the {\thL} variables.
First, given the two eroding bodies $\gamma_i$ and $\gamma_j$ with
lengths $L_i$ and $L_j$, and centers $\cc_i$ and $\cc_j$, we first check
if
\begin{align}
  \label{eqn:BodyCutoff}
  \|\cc_i - \cc_j\| < \frac{L_i}{2\pi} + \frac{L_j}{2\pi} + 
    \alpha \left(\frac{L_i}{N_\iin} + \frac{L_j}{N_\iin} \right),
\end{align}
where $\alpha /geq 1$ is a free parameter that needs to be chosen.
In this manner, we are checking the proximity of the two circles with
centers $\cc_i$ and $\cc_j$, and with radii $L_i$ and $L_j$, where
$\alpha$.  This criteria allows us to quickly determine bodies that
contain discretization points where the Barycentric quadrature rule
should be applied.  Assuming that the two bodies $\gamma_i$ and
$\gamma_j$ satisfy condition~\eqref{eqn:BodyCutoff}, for each point $\xx
\in \gamma_j$, we check if
\begin{align}
  \label{eqn:PointsCutoff}
  \|\xx - \cc_i\| < \frac{L_i}{2\pi} + \alpha_\iin 
    \left(\frac{L_i}{N_\iin} \right).
\end{align}
By applying this method, we do not have to do an all-to-all comparison
of the discretization points on two eroding bodies.

To determine an appropriate value for $\alpha_\iin$, we first compute an
accurate solution by using the Barycentric quadrature rule for all
discretization points.  Then, for multiple values of $\alpha_\iin$, we
compute the velocity field with the trapezoid rule for all points that
do not satisfy~\eqref{eqn:PointsCutoff}.  By comparing these two
velocities, we find that a value of $\alpha_\iin = ???$ gives sufficient
accuracy while keeping the number of points that require the expensive
Barycentric quadrature rule to a minimum.

Given an eroding body $\gamma_i$, to determine if a target point $\xx$
that is in the fluid bulk rather than on another eroding body requires
the Barycentric quadrature rule, we only check
condition~\eqref{eqn:PointCutoff}.  Finally, to determine if points on
$\gamma_i$ are too close to the outer wall, we make use of the fact that
the outer wall is bounded in $[-3,3] \times [-1,1]$, and that the bodies
are all contained in $[-1,1] \times [-1,1]$.  Therefore, we first check
if
\begin{align}
  \|\cc_i \pm 1\| < L_i + \alpha_\out \frac{L_\out}{N_\out}.
\end{align}
If body $\gamma_i$ satisfies this condition, for each point $\xx \in
\gamma_i$, we apply the Barycentric rule to points that satisfy
\begin{align}
  \|\xx \pm 1\| < \alpha_\out \frac{L_\out}{N_\out}.
\end{align}
Using the same strategy as we used for body-body interactions, we find
that an appropriate safety parameter is $\alpha_\out = ???$.

@@

%{\color{red}
%By using Barycentric quadrature rule to approach integrations, we get a reasonable(?) double-layer potential and quantities of interest at target points but it requires $\mathcal{O}(N^2)$ operations per target point which is computational expensive if we impose it on all the target points. In order to accelerate the process and keep the accuracy, it is a compromise that we use Barycentric quadrature rule to approach the nearly singular integratons but apply trapezoid rule when the target point is far from source points. 
%
%
%To numerically compute the completed double-layer potential, we discretize each of the $M$ bodies with  $N_\iin$ points, and the outer wall with $N_\out$ points. The following is the four-step fast summation methods.
%}
%\begin{itemize}
%  \item Apply trapezoid rule as a first pass
%
%{\color{red} We use the Fast Multipole Method (FMM)~\cite{gre-rok1987, gre-gre-may1992} with trapezoid rule to calculate the double-layer potential.
%}
%  \item Efficiently find points that are too close for trapezoid
%
%{\color{red}
% Let $L_w$ be the length of the outer wall and $L_{b}$ be the length of the boundary of body where $b=1, \cdots, M$. Here are the criteria to see if two points are too close for trapezoid. For criterion of outer wall to body,  let $d$ be the minimum distance between outer wall and a point $x$ on the body $b$, we say $x$ is too close to outer wall if
%\begin{equation}\label{eqn:Cw2b}
%d< \alpha_1\frac{L_w}{N_\out}
%\end{equation}
%where $\alpha_1$ is a constant. 
%For criterion between two bodies, let $d$ be the distance between the centroid $a$ of the body $b_1$ and a point $x$ on the body $b_2$, we say they are too close if
%\begin{equation}\label{eqn:Cb2b}
%d<\alpha_2\left(\frac{L_{b_1}}{N_\iin}+\frac{L_{b_2}}{N_\iin}\right)
%\end{equation}
%where $\alpha_2$ is a constant. 
%
%}
%\end{itemize}
%{\color{red}  (Discussion of the choices of $\alpha_1$ and $\alpha_2$ with respect to accuracy and  efficiency)
%
%To find the suitable values of $\alpha_1$ and $\alpha_2$ in~\eqref{eqn:Cw2b} and \eqref{eqn:Cb2b}, we set $\alpha_1=\alpha_2=1, 2, 3$, and $4$ to test the criteria on a 20-body problem with $N_\out = 1024$ and $N_\iin = 128, 256$, and $512$. To check the accuracy of the simulations, we use the solution calculated by \todo[inline]{what method} to \todo[inline]{...}
%To see how efficient our code is, we record the cpu time for each step.
%
%}
